---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Code for loading the required libraries and the data.  
```{r}
## Loading libraries
library(readr)
library(plyr)
library(dplyr)
library(Hmisc)

## Reading the data
activity <- read_csv("activity.zip")
```

2. Processing the data(including all the procesing for further analyses)  
```{r}
a1 <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
a2 <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```

## What is mean total number of steps taken per day?
1. Following is the histogram for the total number of steps taken everyday  
```{r}
hist(a1)
```
The histogram shows the frequency of the total number of steps taken everyday  
2. Followng is the mean and median of the total number of steps taken everyday  
```{r}
mean(a1)
median(a1)
```

## What is the average daily activity pattern?
1. Following is the code for the time series plot of the total number of steps taken everyday  
```{r}
plot(a2, type = "l", lwd = 2)
```

2.The 5-minute interval which has highest avergave on all days is  
```{r}
which.max(a2)
```

## Imputing missing values
1.The total number of missing values in the dataset are:  
```{r}
sum(is.na(activity$steps))
```

2. The missing values are replaced with the mean of the corresponding 5-minute interval  
```{r}
activity2 <- activity %>%
        group_by(interval) %>%
        mutate_each(funs(replace(., which(is.na(.)), mean(., na.rm=TRUE))), 
                                       +                   starts_with('steps'))
```

3. A new dataset called activity2 is created after imputing the missing values  

4. Histogram of the total number of steps taken everyday after imputing the values is as follows:  
```{r}
a3 <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
hist(a3)
```
 The mean and median of the total steps taken everyday in the new dataset is as follows:  
```{r}
mean(a3)
median(a3)
```

## Are there differences in activity patterns between weekdays and weekends?
1. A new factor variable is created with two levels : Weekdays & weekends  
```{r}
weekday1 <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
activity2 <- mutate(activity2, holiday = factor((weekdays(date) %in% weekday1), 
                                                levels=c(FALSE, TRUE), labels=c('weekend', 'weekday') ))
```

2. Plotting the average of the steps taken in each 5-minute interval according to the factor variable holiday:  
```{r}
act2 <- group_by(activity2, holiday, interval) %>% summarise(steps = mean(steps))
ggplot(na.omit(act2), aes(x = interval, y = steps)) + geom_line() + facet_wrap(.~holiday, nrow = 2)
```